A matrix-free approach for finite-strain hyperelastic problems using geometric multigrid

### 2018-09-18 doc/large_strain.tex
section Bibliography
a few more GMG references

Efficient and scalable matrix-vector products is not enough to obtain a method with linear scaling with respect to the number of unknown DoFs $N$.
Efficiency of standard iterative solvers deteriorate for large systems of linear equations.
It has been widely accepted that the geometric multigrid (GMG) preconditioners \cite{Bramble1990, Briggs2000, Janssen2011,May2015} can greatly improve the performance of iterative solvers by smoothing the defect on a hierarchy of levels.
As a result the number of iterations for convergence becomes independent of the number of mesh refinements and thus leads to the $\mathcal O (N)$ solution method. For further information about geometric multigrid methods we refer the reader to \cite{Briggs2000}.

For this work we adopt the matrix-free level operators within the GMG and Chebyshev pre- and post-smoothing on each level. Given a generally heterogeneous non-linear material, definition of transfer operations (restriction and prolongation) and level operators is not trivial. One approach is to start from an arbitrary heterogeneous material at the fine scale and employ homogenization theory \cite{Suquet1985, Hill1972,Hashin1983,Castaneda1997} to design specific transfer operators \cite{Miehe2007}\footnote{
  We note that this approach is limited to the case where for each quadrature point within an element the fourth order tangent operator is the same.
}.
In this approach the fine-scale displacement is decomposed into long-wave and short-wave contributions by additive split, where the former is associated to the homogeneous response of a patch of elements and the latter represents fluctuations due to the heterogeneous structure of the patch. In the case of a homogeneous patch, the transfer operators reduce to the standard geometric transfer. Based on this observation we adopt the following approach. It is assumed that the mesh at the coarsest level can accurately describe the heterogeneous fine-strain elastic material. In this case the restriction and prolongation operations for patches of elements are always performed for the same material and thus admits usage of standard geometric transfer operators. We define the level operators as follows. Recall that the fine scale tangent is obtain by linearization around the current displacement field $\gz u^h$. We then restrict this displacement field to multigrid levels $\{l\}$ and evaluate tangent operators using the matrix-free algorithms from Section \ref{sec:mf}. Essentially level operators correspond to the linearization around smoothed representation of displacement field.
Note that by doing so the operator (tangent) $K^{l+1}$ on level $l+1$ is not directly related to the operator (tangent) $K^{l}$ on level $l$ via $K^{l+1}=I^{l+1}_{l}  K^l I^l_{l+1}$, where $I^l_{l+1}$ and $I^{l+1}_l$ are prolongation (coarse-to-fine) and restriction (fine-to-coarse) operators, respectively.
Finally, at the coarsest level we adopt CG solver which shall reduce residual by three orders of magnitude within at most hundred of steps.
The classical multigrid V-cycle approach with two pre- and post-smoothing is adopted.

@@

>>>>

Efficient and scalable matrix-vector products is not enough to obtain a method with linear scaling with respect to the number of unknown DoFs $N$.
Efficiency of standard iterative solvers deteriorate for large systems of linear equations.
It has been widely accepted that the geometric multigrid (GMG) preconditioners \cite{Bramble1990, Briggs2000, Janssen2011,May2015} can greatly improve the performance of iterative solvers by smoothing the defect on a hierarchy of levels.
As a result the number of iterations for convergence becomes independent of the number of mesh refinements and thus leads to the $\mathcal O (N)$ solution method. For further information about geometric multigrid methods we refer the reader to \cite{Briggs2000,Hackbusch1985,Wesseling1992}.

For this work we adopt the matrix-free level operators within the GMG and Chebyshev pre- and post-smoothing on each level. Given a generally heterogeneous non-linear material, definition of transfer operations (restriction and prolongation) and level operators is not trivial. One approach is to start from an arbitrary heterogeneous material at the fine scale and employ homogenization theory \cite{Suquet1985, Hill1972,Hashin1983,Castaneda1997} to design specific transfer operators \cite{Miehe2007}\footnote{
  We note that this approach is limited to the case where for each quadrature point within an element the fourth order tangent operator is the same.
}.
In this approach the fine-scale displacement is decomposed into long-wave and short-wave contributions by additive split, where the former is associated to the homogeneous response of a patch of elements and the latter represents fluctuations due to the heterogeneous structure of the patch. In the case of a homogeneous patch, the transfer operators reduce to the standard geometric transfer. Based on this observation we adopt the following approach. It is assumed that the mesh at the coarsest level can accurately describe the heterogeneous fine-strain elastic material. In this case the restriction and prolongation operations for patches of elements are always performed for the same material and thus admits usage of standard geometric transfer operators. We define the level operators as follows. Recall that the fine scale tangent is obtain by linearization around the current displacement field $\gz u^h$. We then restrict this displacement field to multigrid levels $\{l\}$ and evaluate tangent operators using the matrix-free algorithms from Section \ref{sec:mf}. Essentially level operators correspond to the linearization around smoothed representation of displacement field.
Note that by doing so the operator (tangent) $K^{l+1}$ on level $l+1$ is not directly related to the operator (tangent) $K^{l}$ on level $l$ via $K^{l+1}=I^{l+1}_{l}  K^l I^l_{l+1}$, where $I^l_{l+1}$ and $I^{l+1}_l$ are prolongation (coarse-to-fine) and restriction (fine-to-coarse) operators, respectively.
Finally, at the coarsest level we adopt CG solver which shall reduce residual by three orders of magnitude within at most hundred of steps.
The classical multigrid V-cycle approach with two pre- and post-smoothing is adopted. For other approaches that combine homogenization and multigrid methods for small strain elastic materials, see \cite{Bayreuther2003,Fish1995}.

@@

\section*{Bibliography}
